A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-06-18 demystify.tex
updates

% \section{How long do you evaluate an application?}\label{sec:ievaluate}

% On average, I spend about 15 minutes reviewing each application (less for clear rejections and more for potential acceptances). In earlier years I spent more time but now  I’ve become more efficient in evaluating..

% The review system compiles a single PDF per applicant, including a summary page, transcripts, test scores, LoRs, a CV, SOP, and any writing samples. I review these documents in order, taking notes to support my evaluations.

% I start with the summary, checking for low GPAs or test scores below the university minumum. I then skim the transcripts for low grades in relevant courses, noting issues like "- many low grades in main courses and - unknown international school with good GPA".

% I pay close attention to strong LoRs and skim weaker ones, noting either strong endorsements from well-known professors or generic content. For the CV, I look for publications, research experiences, and notable achievements, noting things like "+ published papers in top venues" or "+ gold medal in an international competition."

% The SOP is read thoroughly unless it's poorly written; a brief skim suffices in those cases. I note whether the SOP is engaging and well-tailored to our program or if it’s generic and uninspired.

% Finally, I make a decision—usually either a direct rejection or an offer of admission with full funding, mentioning potential funding sources. Exceptionally strong candidates are recommended for the University Presidential fellowship. The admissions committee chair reviews all evaluations, typically following the recommendations, but discrepancies are resolved through discussion among reviewers to reach a consensus.

@@

\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
In many cases, professors or research labs put their projects on GitHub. For example, my research group has many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).

@@

In my opinion, these reasons are not strong enough to justify retracting an acceptance. A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, which many universities participate. Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.

However, regardless of the resolution, retracting an acceptance can have ethical implications. When you accept an offer, you are making a commitment to work with that prof, who then might stop looking for other students. So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also takes away the opportunity from other students. 
Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.

>>>>

\section{How long do you evaluate an application?}\label{sec:ievaluate}

%

On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this time
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

For each application, the GMU application system compiles a single PDF including a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples. I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum. I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.  These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good? is IELTS score sufficient?'')

I pay close attention to \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content. For the \emph{CV}, I look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
I read \emph{SOP} thoroughly (unless it's poorly written then I just skim). I note whether the SOP is exciting, research-oriented, and tailored to our program. All these notes are entered into the evaluation system.

Finally, I make a decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA). I also recommend very strong candidates for the University Presidential fellowship. The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus. Typically there is no involvement from the adcom chair, department chair, or anyone else (i.e., the decision is made by the reviewers).

@@

\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
In many cases, professors or research labs put their projects on GitHub. For example, my research group, \href{https://github.com/dynaroars/}{Dynaroars}, has many open-source projects that undergrads can contribute to.

@@

In my opinion, these reasons are not strong enough to justify retracting an acceptance. A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, in which many universities participate. Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.

However, regardless of the resolution, retracting an acceptance can have ethical implications. When you accept an offer, you are committing to work with that prof, who then might stop looking for other students. So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students. 
Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.

### 2024-06-18 faculty.tex
updates

\chapter{Research}
\chapter{Obtaining Funding}
Getting funding is arguably the most important part of getting tenure.  Most new professors struggle with grant writing, which is very different compared to research papers. However, without funding, you can't pay your students, conference registration and travel, or even your own summer salary. In short, the lack of funding causes lost of stress, uncertainty, and threaten your career.  In this chapter, we will discuss the various ways to obtain funding.

\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.  Then eventually they will get used to it and learn to like it (or at least tolerate it).  The key is to keep trying, and to not give up. The best researchers at the best universities get rejected all the time.  It's part of the process.  So don't get discouraged.  Keep trying, and you will get better over time.

@@

It's a very different skill than writing research papers.  However, it's a skill that can be learned.  The best way to learn is to read other people's grants, and to write your own.  You will likely get rejected many times, but that's normal.  You will get better over time.  The key is to keep trying, and to not give up.  If you are not getting rejected, you are not aiming high enough.  The best researchers in the world get rejected all the time.  It's part of the process.  So don't get discouraged.  Keep trying, and you will get better over time.
\section{NSF and other Government Funding}

@@

\section{Internal Funding}

\section{Industrial Gifts}

>>>>

\chapter{Research}
\chapter{Obtaining Funding}
Getting funding is arguably the most important part of getting tenure.  Writing grant propsal is very different from writing research papers. However, without funding, you can't pay your students, conference registration and travel, or even your own summer salary. %In short, the lack of funding causes lost of stress, uncertainty, and threaten your career.
Most new professors struggle with grant writing at first, but eventually get better at it.  In this chapter, we will discuss the various ways to obtain funding.

%Like most new professors, I struggled big time grant writing at first, but eventually got better at it.

%, which is very different compared to research papers.   In this chapter, we will discuss the various ways to obtain funding.

\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.  Then eventually they will get used to it and learn to like it (or at least do not hate it as much).  So don't get discouraged.  Keep trying, and you will get better over time.

@@

It's a very different skill than writing research papers.  However, it's a skill that can be learned.  The best way to learn is to read other people's grants, and to write your own.  You will likely get rejected many times, but that's normal.  You will get better over time.  The key is to keep trying, and to not give up.  If you are not getting rejected, you are not aiming high enough.  The best researchers in the world get rejected all the time.  It's part of the process.  So don't get discouraged.  Keep trying, and you will get better over time.

\section{Types of Funding}
\subsection{NSF and other Government Funding}

@@

\subsection{Internal Funding}

\subsection{Industrial Gifts}

\section{My Story}
